---
title: "Time series"
author: "Swati Gupta"
date: "4/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
rm(list=ls())
library(forecast)
library(vars)
data = read.csv("datanew.csv",header=T)
```

```{r}
temp.ts = ts(as.numeric(data$temperature),start=c(2014),frequency=365.25)
wind.ts = ts(as.numeric(data$wind_speed),start=c(2014),frequency=365.25)
data.ts = ts.union(temp.ts,wind.ts)
par(mfrow=c(2,2))
plot(temp.ts, type="l",main="Temperature",ylab="Temperature")
hist(temp.ts, main="Distribution of values",xlab="Temperature")
acf(temp.ts,main="ACF plot for Temperature")  
pacf(temp.ts,main="PACF plot for Temperature")
par(mfrow=c(2,2))
plot(wind.ts, type="l",main="Wind",ylab="Wind")
hist(wind.ts, main="Distribution of values",xlab="Wind")
acf(wind.ts, main="ACF plot for Wind")  
pacf(wind.ts, main="PACF plot for Wind")
```
Looking at the data we see clear annual seasonality. To account for that we difference the data. There is also variable variance in data with time. To account for this the log of the data will be taken

```{r}
dtemp=diff(temp.ts, lag=365)
dwind=diff(wind.ts,lag = 365)
data.ts=ts.union(dtemp,dwind)
plot(data.ts, type="l",main="")
acf(data.ts)
pacf(data.ts)
```
Now that we have stationary data, we now dicide it into training and testing to try different models

```{r}
data=data.ts
n = nrow(data)
data.train=data[1:(n-114),]
data.test=data[(n-113):n,]
temp_train=ts(dtemp[1:(n-114)],start=c(2014),frequency=365.25)
wind_train=ts(dwind[1:(n-114)],start=c(2014),frequency=365.25)
temp_test=ts(dtemp[(n-113):n],start=c(2019),frequency=365.25)
wind_test=ts(dwind[(n-113):n],start=c(2019),frequency=365.25)
```


We now try to fit an ARIMA model to temperature. We do model selection based on AIC

```{r eval=FALSE, include=FALSE}
#### Univariate ARIMA model

final.aic = Inf
final.order = c(0,0,0)
for (p in 11:13) for (d in 0:2) for (q in 7:9) {
  #print(c(p, d, q))
   current.aic = AIC(arima(temp_train, order=c(p, d, q), method="ML"))
   if (current.aic < final.aic) {
     final.aic = current.aic
     final.order = c(p, d, q)
     
   }
 }
final.order
# [1] 12  1  8
```


```{r}
model.arima = arima(temp_train, order=final.order, method="ML")
summary(model.arima)
```
```{r}
## Residual analysis
par(mfrow=c(2,2))
plot(resid(model.arima), ylab='Residuals',type='o',main="Residual Plot")
abline(h=0)
acf(resid(model.arima),main="ACF: Residuals")
hist(resid(model.arima),xlab='Residuals',main='Histogram: Residuals')
qqnorm(resid(model.arima),ylab="Sample Q",xlab="Theoretical Q")
qqline(resid(model.arima))
```



```{r}
fore = forecast(model.arima,h=114)
fore=as.data.frame(fore)
point.fore = ts(fore[,1],start=c(2019),frequency=365.25)
lo.fore = ts(fore[,4],start=c(2019),frequency=365.25)
up.fore = ts(fore[,5],start=c(2019),frequency=365.25)
ymin=min(c(temp_train[(n-478):n],lo.fore))
ymax=max(c(temp_train[(n-478):n],up.fore))
plot(ts(dtemp[(n-478):n],start=c(2018),frequency=365.25),  ylab="Differenced Temperature", type="l",main="Predictions from ARIMA model")
points(point.fore,lwd=2,col="red")
lines(lo.fore,lty=3,lwd= 2, col="blue")
lines(up.fore,lty=3,lwd= 2, col="blue")
legend("topleft",c("Actual value","Estimated value","95% interval"), bty="n", lwd=c(1,2,2), col=c("black","red","blue"))
```

```{r}
VARselect(data.train, lag.max = 20,type="both")$selection
```

```{r}
model.var=VAR(data.train, p=8,type="both")
summary(model.var)
```

```{r}
pred = predict(model.var, n.ahead=114, ci=0.95)[[1]]$dtemp
point.pred = ts(pred[,1],start=2019, freq=365.25)
lo.pred = ts(pred[,2],start=2019, freq=365.25)
up.pred = ts(pred[,3],start=2019, freq=365.25)

plot(ts(dtemp[(n-478):n],start=c(2018),frequency=365.25),  ylab="Differenced Temperature", type="l",main="Predictions from VAR Model")
points(point.pred,lwd=2,col="red")
lines(lo.pred,lty=3,lwd= 2, col="blue")
lines(up.pred,lty=3,lwd= 2, col="blue")
legend("topleft",c("Actual value","Estimated value","95% interval"), bty="n", lwd=c(1,2,2), col=c("black","red","blue"))
```

